---
title: "AVP Ticket Sales:Fall 2016, Spring 2017"
author: "Shade Wilson"
date: "8/9/2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(lubridate)
library(stringr)

theme_set(theme_grey(base_size = 18)) 
```

```{r readin, include = FALSE}
# read in data
fall_read <- read_csv(file = "~/Desktop/Coding/R/AVP_ticket_sales/AVP_Fall2016_Ticket_Sales.csv",
                      col_types = cols(Date = col_character(),
                                       `Time (roughly) (HH:MM)` = col_character()))
spring_read <- read_csv(file = "~/Desktop/Coding/R/AVP_ticket_sales/AVP_Spring2017_Ticket_Sales.csv",
                        col_types = cols(Date = col_character(),
                                         `Time (roughly) (HH:MM)` = col_character()))
```

```{r rename, include = FALSE}
# Rename variable names so they 1) look better/more concise and 2) fit on screen better
# add new, useful variables
fall <- fall_read %>% 
  mutate(date = mdy(Date),
         wday = wday(date, label = TRUE),
         date_time = mdy_hm(str_c(Date, `Time (roughly) (HH:MM)`)),
         fri_stud = `Friday tickets Sold`,
         sat_stud = `Sat tickets Sold`,
         fri_deal =  `Friday Deal Tickets Sold`,
         fri_adult = `Friday Adult tickets`,
         sat_adult = `Sat Adult Tickets`,
         method = `Payment Method (cash, card, venmo, online)`,
         total = `Total Tix`,
         profit = (fri_stud + sat_stud) * 6 + (fri_adult + sat_adult) * 10 + fri_deal * 12/3) %>% 
  select(wday:profit)

spring <- spring_read %>% 
  mutate(date = mdy(Date),
         wday = wday(date, label = TRUE),
         date_time = mdy_hm(str_c(Date, `Time (roughly) (HH:MM)`, sep  = " ")),
         fri_stud = `Friday tickets Sold`,
         sat_stud = `Sat tickets Sold`,
         fri_adult = `Friday Adult tickets`,
         sat_adult = `Sat Adult Tickets`,
         method = `Payment Method (cash, card, venmo, online)`,
         total = `Total Tix`,
         profit = (fri_stud + sat_stud) * 7 + (fri_adult + sat_adult) * 12) %>% 
  select(wday:profit)
```

Example data frame (from fall 2016), extrapolated from data collected: date and time purchased, 
student or adult, which night, and payment method. Very easy to keep track of. So far we have
(semi) detailed sales records for two semesters, and it's important to continue with this trend
so AVP can make the most informed decisions possible.
```{r print, echo = FALSE}
print(fall, n = 10, width = Inf)

```

### Total Tickets Sold Per Night and Per Category

```{r sum_cols, include = FALSE}
sum_cols <- function(df) {
  output <- vector("integer", ncol(df))
  for (i in seq_along(df)) {
    if (is.numeric(df[[i]])) {
      output[[i]] <- sum(df[[i]])
    }
  }
  output <- set_names(output, names(df))
  output <- output[!output == 0]
}
fall_totals <- sum_cols(fall)
spring_totals <- sum_cols(spring)
```

```{r profit_by_tix, include = FALSE}
profit_by_tix <- function(vector, col_names, prices) {
  indiv_profit <- vector("double", length(col_names))
  if (length(col_names) != length(prices)) {
    stop("'col_names' and 'prices' must be the same length", call. = FALSE)
  }
  prices <- set_names(prices, col_names)
  for (name in col_names) {
    indiv_profit[name] <- vector[name] * as.numeric(prices[name])
  }
  sum(indiv_profit)
}

fall_fri_names <- c("fri_stud", "fri_deal", "fri_adult")
fall_sat_names <- c("sat_stud", "sat_adult")
spring_fri_names <- c("fri_stud", "fri_adult")
spring_sat_names <- c("sat_stud", "sat_adult")

# student, deal, and adult prices
fall_fri_prices <- c(6, 12/3, 10)
fall_sat_prices <- c(6, 10) # excluding deal
spring_prices <- c(7, 12)

fall_costs <- 500 + 38 + 11
spring_costs <- 550 + 55/2 + 150/2
```


```{r profit, include = FALSE}
fall_profit <- list("fall 2016", sum(fall_totals[fall_fri_names]),
                 sum(fall_totals[fall_sat_names]),
                 fall_totals["total"], 
                 profit_by_tix(fall_totals, col_names = fall_fri_names, prices = fall_fri_prices),
                 profit_by_tix(fall_totals, col_names = fall_sat_names, prices = fall_sat_prices),
                 fall_costs)
fall_profit <- set_names(fall_profit, c("concert", "friday", "saturday", "total", "fri_profit", 
                                        "sat_profit", "costs"))
spring_profit <- list("spring 2017", sum(spring_totals[spring_fri_names]),
                 sum(spring_totals[spring_sat_names]),
                 spring_totals["total"],
                 profit_by_tix(spring_totals, 
                               col_names = spring_fri_names, 
                               prices = spring_prices),
                 profit_by_tix(spring_totals, 
                               col_names = spring_sat_names, 
                               prices = spring_prices),
                 spring_costs)
spring_profit <- set_names(spring_profit, c("concert", "friday", "saturday", "total", "fri_profit", 
                                            "sat_profit", "cost"))

both_sum <- rbind(fall_profit, spring_profit) %>% 
  as.tibble() %>% 
  unnest() %>% 
  gather(friday, saturday, key = "night", value = "tickets") %>% 
  mutate(gross = ifelse(night == "friday", fri_profit, sat_profit),
         net = gross - costs) %>% 
  select(-total, -fri_profit, -sat_profit) %>% # deselect unnecessary columns
  select(concert, night:gross, costs, net) # reorder (move costs to last column)
```

Despite a 3-for-2 deal on Friday in the fall, more tickets were sold in the spring overall, and the
gross profit from the spring eclispes that of the fall. In the fall, `r both_sum$tickets[1]` tickets
were sold on Friday and `r both_sum$tickets[3]` on Saturday for a total of 
`r both_sum$tickets[1] + both_sum$tickets[3]` tickets overall. In the spring,  sales reached
`r both_sum$tickets[2]` tickets on Friday and `r both_sum$tickets[4]` on Saturday, 
bringing the total to `r both_sum$tickets[2] + both_sum$tickets[4]` tickets sold.

Ticket sales in the fall grossed  $`r both_sum$gross[1] + both_sum$gross[3]` and 
$`r both_sum$gross[2] + both_sum$gross[4]` in the spring. The costs per night in the fall
were as follows: $500 for Jerry, $38 for awesome, $22 for security guard (one night). Additionally,
were was a cost for printing programs, but the records have been lost. More likely, they never
existed. In the spring, the costs were: $1100 for Jerry, $55 for awesome, and $150 for the 
security guard(s).

Figuring in the total costs, the gross profit for the fall was $`r both_sum$net[1] + both_sum$net[3]`
and was $`r both_sum$net[2] + both_sum$net[4]` in the spring. The total net profits for the 2016-2017
year were $`r sum(both_sum$net)`. At this rate, it would take eight more concerts to pay for a 
$10,000 CD.




```{r ticket_sold, echo = FALSE, out.width = '33%'}
ggplot(both_sum, aes(night, tickets, fill = concert)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "",
       y = "",
       title = "Tickets Sold") +
  theme(legend.position="none")

ggplot(both_sum, aes(night, gross, fill = concert)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "",
       y = "",
       title = "Gross Profit") +
  theme(legend.position="none")

ggplot(both_sum, aes(night, net, fill = concert)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "",
       y = "",
       title = "Net Profit")

```


### Payment Methods

Around a third of total tickets for both semesters were sold at the door, so for the future, 
don't lose hope if tabling and online sales aren't exactly what you want. Door sales may be
influenced by the weather and the amount of time into the semester: the better the weather and
the earlier into the semester, the more likely people are to go to an a cappella concert on a whim.


```{r payment_methods, echo = FALSE, out.width="50%"}
fall_methods <- fall %>% 
  group_by(method) %>% 
  count()
spring_methods <- spring %>% 
  filter(!is.na(method)) %>% 
  group_by(method) %>% 
  count()

pay_methods <- left_join(fall_methods, spring_methods, by = "method") %>% 
  mutate(fall = n.x, 
         spring = n.y) %>% 
  select(-n.x, -n.y)

pay_methods %>% 
  gather(fall, spring, key = "concert", value = "n") %>% 
  ggplot(aes(reorder(method, n, FUN = max), n, fill = concert)) +
    geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Payment Method",
       y = "") +
  theme(legend.position="none")

pay_methods %>% 
  mutate(simple_method = if_else(method == "door", "door", "other")) %>% 
  group_by(simple_method) %>% 
  summarise(fall = sum(fall),
            spring = sum(spring)) %>% 
  gather(fall, spring, key = "concert", value = "n") %>% 
  ggplot(aes(reorder(simple_method, n, FUN = max), n, fill = concert)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Payment Method",
       y = "")
```


### Ticket Sales per Day

```{r sales_perday_setup}
fall_by_day <- fall %>% 
  group_by(wday) %>% 
  summarize(n = sum(total),
            profit = sum(profit)) %>% 
  mutate(cum_profit = cumsum(profit))
spring_by_day <- spring %>% 
  group_by(wday) %>% 
  summarize(n = sum(total),
            profit = sum(profit)) %>% 
  filter(!is.na(n)) %>% 
  mutate(cum_profit = cumsum(profit))

both_sales <- left_join(fall_by_day, spring_by_day, by = "wday") %>% 
  mutate(fall = n.x, 
         spring = n.y) %>% 
  gather(fall, spring, key = "concert", value = "n") %>%
  mutate(profit = ifelse(concert == "fall", profit.x, profit.y),
         profit = ifelse(is.na(profit), 0, profit),
         n = ifelse(is.na(n), 0, n),
         cum_profit = ifelse(concert == "fall", cum_profit.x, cum_profit.y),
         cum_profit = ifelse(is.na(cum_profit), 0, cum_profit)) %>% 
  select(-(n.x:cum_profit.y))

s <- both_sales %>% 
  ggplot(aes(wday, profit, fill = concert)) +
    geom_bar(stat = "identity", position = "dodge") +
    labs(x = "Day of the Week",
         y = "Gross Profit")
```

Sales don't seem to be significant until after Wednesday. This isn't surprising since people
tend to procrastinate on buying tickets and also wait until later in the week to solidify weekend
plans. The most profit was made on Friday even though Saturday pulled in more profit 
both semesters by far because tickets for both nights could be purchased then.

```{r sales_perday, out.width="50%"}
s + theme(legend.position="none")
s + geom_bar(aes(wday, cum_profit, fill = concert),
             stat = "identity", position = "dodge", alpha = 1/3) +
  labs(title = "Cummulative Sales and Sales by Day")
```


### Tabling Ticket Sales

Sales peak around noon (lunchtime). Expect sales to be particularly slow on Tuesdays.
Ticket sales before 11 are few and far between. Due to the 
hecticness of the week resulting in lack of sleep, I personally don't believe it's worth harrassing 
AVP members out of bed early to sell tickets to one person. However, the table should certainly
be up and running before 11 to catch those leaving their 10 AMs/9:30s. Consistency is the most 
important thing to keep in mind. Tabling is very weather dependent and success varies wildly, so
the data presented here certainly cannot be considered definitive.

Overall, tickets sold at the table are almost negligible, but the advertising is not. Keep it in
mind that the most important part of tabling is getting word of the concert out. The only people 
that really buy tickets at the table are those that know a lot about AVP and are sure they're coming
before the week even begins (friends and that one groupie) and those that buy a ticket on a whim
(very few in my experience). I would recommend printing advert fliers to hand out to the overwhelming
majority of students who have no idea about the group or are on the fence.

```{r tabling_setup}
fall_lims <- c(ymd_hm("2016-01-01 10:00","2016-01-01 15:00"))
spring_lims <- c(ymd_hm("2017-01-01 10:00","2017-01-01 15:00"))
  
fall_tabling <- fall %>% 
  filter(!wday %in% c("Sun", "Sat")) %>% 
  mutate(date_fixed = update(date_time, yday = 1)) %>% 
  ggplot(aes(date_fixed)) +
  geom_histogram(fill="lightblue",
                 color = "grey50",
                 binwidth = 15*60) +
  labs(x = "Time (hour)", y = "Count",
       title = "Fall 2016 Tabling") +
  scale_x_datetime(limits = fall_lims, date_labels = "%I")

spring_tabling <- spring %>% 
  filter(!wday %in% c("Sun", "Sat")) %>% 
  mutate(date_fixed = update(date_time, yday = 1)) %>% 
  ggplot(aes(date_fixed)) +
  geom_histogram(fill="lightblue",
                 color = "grey50",
                 binwidth = 15*60) +
  labs(x = "Time (hour)", y = "Count",
       title = "Spring 2017 Tabling") +
  scale_x_datetime(limits = spring_lims, date_labels = "%I")
```


```{r tabling, out.width="50%", warning=FALSE}
fall_tabling
spring_tabling

fall_tabling + facet_wrap(~wday)
spring_tabling + facet_wrap(~wday)
```

